---
title: "Probability code exercise with R along with questions"
author: "Amirali Satvati"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    toc_depth: 2
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1

A coin is tossed 5 times. Let \( X \) represent the number of observed heads. Calculate the following probabilities:

## 1.1 The probability of observing exactly 4 heads:

```{r}
dbinom(4, size = 5, prob = 0.5)
```

## 1.2 The probability of observing at least 2 heads:

```{r}
1 - pbinom(1, size = 5, prob = 0.5)
```

## 1.3 The probability of observing no heads:

```{r}
dbinom(0, size = 5, prob = 0.5)
```

# Exercise 2

A die is rolled twice. Let \( X_1 \) and \( X_2 \) represent the outcomes of the first and second rolls, respectively. Define \( Y = \min(X_1, X_2) \). Calculate the following probabilities:

## 2.1 The probability that \( Y < 5 \):

```{r}
mean(replicate(10000, {
  min(sample(1:6, size = 1, replace = TRUE), sample(1:6, size = 1, replace = TRUE)) < 5
}))
```

## 2.2 The probability that \( Y = 2 \):

```{r}
mean(replicate(10000, {
  min(sample(1:6, size = 1, replace = TRUE), sample(1:6, size = 1, replace = TRUE)) == 2
}))
```

## 2.3 The probability that \( Y \leq 3 \):

```{r}
mean(replicate(10000, {
  min(sample(1:6, size = 1, replace = TRUE), sample(1:6, size = 1, replace = TRUE)) <= 3
}))
```

# Exercise 3

A bag contains 5 white marbles, 7 green marbles, and 12 blue marbles. Calculate the following probabilities:

## 3.1 The probability that if 2 marbles are drawn (with replacement), both are white:

```{r}
wn <- 5
gn <- 7
bn <- 12
nn <- wn + gn + bn

p_both_white <- choose(wn, 2) / choose(nn, 2)
p_both_white
```

## 3.2 The probability that if 3 marbles are drawn (with replacement), their colors are not identical:

```{r}
p_not_identical <- 1 - ((wn / nn)^3 + (gn / nn)^3 + (bn / nn)^3)
p_not_identical
```