---
title: "Pencil Dataset Analysis"
author: "Amirali Satvati"
date: "2025-06-04"
output:
  html_document:
    css: "mystyle1.css"      
    toc: true                
    toc_float: true          
    highlight: "zenburn"     
    theme: "flatly"          
editor_options:
  chunk_output_type: console  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, digits = 2)

knitr::knit_hooks$set(inline = function(x) {
  if (is.numeric(x)) round(x, 2) else x
})
```
# Loading Data
```{r}
library(car)
library(DescTools)
```

# Question 1
Is the mean Sepal Length the same across all three species?
```{r}
summary(aov(Sepal.Length ~ Species, data = iris))
```
Since the p-value is small, the null hypothesis is rejected; thus, the mean Sepal Length differs across the three species.

# Question 2
Is the mean Sepal Length the same for setosa and virginica?
```{r}
data2 <- subset(iris, Species %in% c("setosa", "virginica"))
t.test(Sepal.Length ~ Species, data = data2)
```
Since the p-value is small, the null hypothesis is rejected; the means differ significantly between the two species.

# Question 3
Is the mean Petal Width the same across all three species?
```{r}
summary(aov(Petal.Width ~ Species, data = iris))
```
The p-value is small; thus, the mean Petal Width differs significantly across the three species.

# Question 4
Is the variance of Sepal Length the same for setosa and versicolor?
```{r}
data3 <- subset(iris, Species %in% c("setosa", "versicolor"))
var.test(Sepal.Length ~ Species, data = data3)
```
Since p < 0.05, the null hypothesis is rejected; the variance differs between the two species.

# Question 5
Is the variance of Sepal Length the same across all three populations?
```{r}
leveneTest(Sepal.Length ~ Species, data = iris, center = mean)
```
The p-value is small; thus, the variances are not equal across the three species.

# Question 6
Is the variance of Petal Width the same across all three populations?
```{r}
leveneTest(Petal.Width ~ Species, data = iris, center = mean)
```
Result: The variance of Petal Width is not the same across the three iris species.
